Dynamics of Colombo's Top: Generating Exoplanet Obliquties from Planet-Disc Interactions

### 2020-04-13 writeups/Su_disk_disp.tex
add theta_f definition

We start from Eq.~\eqref{eq:dsdt_base}, which governs the evolution of the spin
axis. We choose coordinate axes such that $\uv{l} = \uv{z}$ and $ \uv{l}_{\rm d}
= \uv{z} \cos I + \uv{x}\sin I$, giving
\begin{equation}
    \rd{\uv{s}}{\tau} = \s{
        \p{\eta \cos I - 1}\uv{z} + \eta \sin I \,\uv{x}} \times \uv{s}.
\end{equation}
Define $S = {\hat s}_{\rm x} + i{\hat s}_{\rm y}$, we find
\begin{equation}
    \rd{S}{\tau} = i\p{\eta\cos I - 1}S
        - i \eta \sin I\cos\theta.\label{eq:nonad_ode}
\end{equation}
To proceed, we assume the obliquity is roughly constant, $\cos \theta \approx
\cos \theta_{\rm i}$. Eq.~\eqref{eq:nonad_ode} can then be solved explicitly,
starting from the initial value $S(\tau_{\rm i})$:
\begin{equation}
    S(\tau)e^{-i\Phi(\tau)} - S(\tau_{\rm i})
        \simeq -i\sin I \cos \theta_{\rm i}
            \int_{\tau_{\rm i}}^\tau \eta(\tau')
            e^{-i\Phi(\tau')}\;\mathrm{d}\tau',
\end{equation}
where
\begin{equation}
    \Phi(\tau) \equiv \int_{\tau_{\rm i}}^\tau \p{\eta(\tau') \cos I
        - 1}\; \mathrm{d}\tau'.
\end{equation}
We now invoke the stationary phase approximation, so that $\Phi(\tau)\simeq
\Phi(\tau_0) + (1/2)\ddot\Phi (\tau_0)(\tau - \tau_0)^2$ where $\tau_0$ is
determined by $\dot\Phi = 0$, occuring when $\eta_0 = 1/\cos I$. We then find,
for $\tau \gg \tau_0$,
\begin{equation}
    S(\tau)e^{-i\Phi(\tau)} - S(\tau_{\rm i})
        \simeq -i \eta(\tau_0)\sin I \cos \theta_{\rm i} e^{-i\Phi(\tau_0)}
            \p{\frac{2\pi}{i \ddot\Phi(\tau_0)}}^{1/2}.
\end{equation}
Using $\dot\eta = -\epsilon \eta$ [Eq.~\eqref{eq:deta_dt}] and
$\ddot\Phi(\tau_0) = \dot\eta(\tau_0)\cos I = -\epsilon$, we have
\begin{equation}
    S(\tau)e^{-i\Phi(\tau)} - S(\tau_{\rm i})
        \simeq -i^{3/2}\tan I \cos \theta_{\rm i} e^{-i\Phi(\tau_0)}
            \p{\frac{2\pi}{\epsilon}}^{1/2}.
\end{equation}
If the initial obliquity is much smaller than the final obliquity ($\theta_{\rm
i} = \abs{S(\tau_{\rm i})} \ll \abs{S(\tau_{\rm f})}$), then we obtain
\begin{equation}
    \theta_{\rm f}\simeq \sqrt{\frac{2\pi}{\epsilon}}\tan I\cos \theta_{\rm i}.
        \label{eq:nonad_q_f}
\end{equation}
This expression is valid only if $\cos \theta \approx \cos \theta_{\rm i}$
throughout the evolution. This corresponds to the limit where $\theta_{\rm f}$
is not too much larger than $\theta_{\rm i}$, which requires $\epsilon$ not to
be too small. Numerically, we verify this is consistent with being in the
nonadiabatic regime $\epsilon \gtrsim \epsilon_{\rm c}$ (see
Figs.~\ref{fig:nonad_3_scan} and~\ref{fig:nonad_3_scan_20}).

>>>>

We start from Eq.~\eqref{eq:dsdt_base}, which governs the evolution of the spin
axis. We choose coordinate axes such that $\uv{l} = \uv{z}$ and $ \uv{l}_{\rm d}
= \uv{z} \cos I + \uv{x}\sin I$, giving
\begin{equation}
    \rd{\uv{s}}{\tau} = \s{
        \p{\eta \cos I - 1}\uv{z} + \eta \sin I \,\uv{x}} \times \uv{s}.
\end{equation}
Define $S = {\hat s}_{\rm x} + i{\hat s}_{\rm y}$, we find
\begin{equation}
    \rd{S}{\tau} = i\p{\eta\cos I - 1}S
        - i \eta \sin I\cos\theta.\label{eq:nonad_ode}
\end{equation}
To proceed, we assume the obliquity is roughly constant, $\cos \theta \approx
\cos \theta_{\rm i}$. Eq.~\eqref{eq:nonad_ode} can then be solved explicitly,
starting from the initial value $S(\tau_{\rm i})$:
\begin{equation}
    S(\tau)e^{-i\Phi(\tau)} - S(\tau_{\rm i})
        \simeq -i\sin I \cos \theta_{\rm i}
            \int_{\tau_{\rm i}}^\tau \eta(\tau')
            e^{-i\Phi(\tau')}\;\mathrm{d}\tau',
\end{equation}
where
\begin{equation}
    \Phi(\tau) \equiv \int_{\tau_{\rm i}}^\tau \p{\eta(\tau') \cos I
        - 1}\; \mathrm{d}\tau'.
\end{equation}
We now invoke the stationary phase approximation, so that $\Phi(\tau)\simeq
\Phi(\tau_0) + (1/2)\ddot\Phi (\tau_0)(\tau - \tau_0)^2$ where $\tau_0$ is
determined by $\dot\Phi = 0$, occuring when $\eta_0 = 1/\cos I$. We then find,
for $\tau \gg \tau_0$,
\begin{equation}
    S(\tau)e^{-i\Phi(\tau)} - S(\tau_{\rm i})
        \simeq -i \eta(\tau_0)\sin I \cos \theta_{\rm i} e^{-i\Phi(\tau_0)}
            \p{\frac{2\pi}{i \ddot\Phi(\tau_0)}}^{1/2}.
\end{equation}
Using $\dot\eta = -\epsilon \eta$ [Eq.~\eqref{eq:deta_dt}] and
$\ddot\Phi(\tau_0) = \dot\eta(\tau_0)\cos I = -\epsilon$, we have
\begin{equation}
    S(\tau)e^{-i\Phi(\tau)} - S(\tau_{\rm i})
        \simeq -i^{3/2}\tan I \cos \theta_{\rm i} e^{-i\Phi(\tau_0)}
            \p{\frac{2\pi}{\epsilon}}^{1/2}.
\end{equation}
If the initial obliquity is much smaller than the final obliquity ($\theta_{\rm
i} = \abs{S(\tau_{\rm i})} \ll \theta_{\rm f} = \abs{S(\tau_{\rm f})}$), then we
obtain
\begin{equation}
    \theta_{\rm f}\simeq \sqrt{\frac{2\pi}{\epsilon}}\tan I\cos \theta_{\rm i}.
        \label{eq:nonad_q_f}
\end{equation}
This expression is valid only if $\cos \theta \approx \cos \theta_{\rm i}$
throughout the evolution. This corresponds to the limit where $\theta_{\rm f}$
is not too much larger than $\theta_{\rm i}$, which requires $\epsilon$ not to
be too small. Numerically, we verify this is consistent with being in the
nonadiabatic regime $\epsilon \gtrsim \epsilon_{\rm c}$ (see
Figs.~\ref{fig:nonad_3_scan} and~\ref{fig:nonad_3_scan_20}).
